{\boldmath Pseudoscalar top-Higgs coupling: Exploration of $\mathrm{CP}$-odd observables to resolve the sign ambiguity }

### 2016-03-02 Paper_ttbarH_Draft.tex
KK edits up to Eq. (30)

%
Before turning to a discussion of $\mathrm{CP}$-odd observables,
it is useful to consider a few theoretical aspects of
the process $\ppprocess$.  In this section we derive a ``factorized''
expression for the gluon fusion contribution to this process.
We then use this expression to isolate various mathematical quantities
that will be useful as we construct $\mathrm{CP}$-odd observables.\par

At the LHC $t\bar{t}H$ production proceeds via $q\bar{q}$
annihilation and $gg$ fusion processes. The relevant leading-order Feynman
diagrams are displayed in Fig.~\ref{fig1}, where the first two
rows show the $q\bar{q}$ and $gg$ $s$-channel diagrams, and
the last one depicts the $gg$ $t$-channel diagrams.  Three more
$gg$-initiated diagrams are obtained by exchanging the gluon
lines in the third row.
We describe the $tH$ coupling with the effective Lagrangian
%
\beq
\label{eq1}
\mathcal{L}_{t\bar{t}H}=-\frac{m_t}{v}(\kp \tbar t+i\kpt \tbar
\gamma_5 t)H,
\eeq
%
where $v=246~\mathrm{GeV}$ is the SM Higgs vacuum
expectation value, and the coefficients $\kp$ and $\kpt$ parameterize 
the scalar and pseudoscalar interaction, respectively. The
SM case is obtained for $\kp=1$ and $\kpt=0$, while the values $\kp=0$
and $\kpt\neq 0$ parameterize a $\mathrm{CP}$-odd Higgs boson. \par
%%%%%%%%%%%%%%%%%%%% Figura 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/home/nico/jaxodraw-2.1-0/
\begin{center}
\begin{figure}[H]
\centering
%\hspace*{-0.4cm}
\subfloat{\includegraphics[scale=0.45]{qqs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1a}}
\subfloat{\includegraphics[scale=0.45]{qqs2_II.pdf}}
%\label{fig1b}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1b}}
\subfloat{\includegraphics[scale=0.45]{ggs2_II.pdf}}
%\label{fig1c}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggt1_II.pdf}}
%\label{fig1b}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt2_II.pdf}}
%\label{fig1c}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt3_II.pdf}}
%\label{fig1b}}
\vspace*{0.02\textwidth}
\caption{Tree-level Feynman diagrams contributing to $t\tbar H$ production
  at the LHC. Three more diagrams are obtained by exchanging the gluon
  lines in the $t$-channel diagrams.}
\label{fig1}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Factorized expression for the scattering cross section}

In this subsection we focus on the gluon fusion contributions
to $t\tbar H$ production, since these dominate over the
the quark-antiquark annihilation contributions.  Furthermore,
we consider the case in which the top and antitop both decay
semileptonically.  As we shall show below, assuming the
narrow width approximation for the top and antitop quarks, the unpolarized
differential cross section for $gg\to t(\to
b{\ell}^+\nu_{\ell})\,\tbar(\to \bbar {\ell}^- \nubar)\,H$ may be
written in the following ``factorized'' form,\footnote{The reader
  is referred to the discussion following Eq.~(\ref{eq17}) for some qualifying
  remarks regarding the ``factorization'' of this expression.}
%
\beq
\label{eq2}
%dd\sigma(gg\to (bl^+\nu_l)(\bbar l^- \nubar) H)=\sum_{\substack{bl^+\nu_l \\ \tiny{\mathrm{spins}}}}\,\sum_{\substack{\bbar l^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,d\Gamma(t(n_t)\to bl^+\nu_l)\,d\Gamma(\tbar (n_{\tbar})\to \bbar l^-\nubar)
d\sigma =\sum_{\substack{b{\ell}^+\nu_l \\ \tiny{\mathrm{spins}}}}\,
   \sum_{\substack{\bbar {\ell}^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,
   d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,
   d\Gamma(t\to b{\ell}^+\nu_{\ell})\,
   d\Gamma(\tbar \to \bbar {\ell}^-\nubar),
\eeq  
%
where $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$ is the differential
cross section for the production of a top and antitop quark,
with spin vectors $n_t$ and $n_{\bar{t}}$, respectively, along with a Higgs
boson.  Also, $d\Gamma(t\to b{\ell}^+\nu_{\ell})$ and
$d\Gamma(\tbar \to \bbar {\ell}^-\nubar)$ are the
partial differential decay widths for an unpolarized top and
anti-top quark.  The four-vectors
$n_t$ and $n_{\tbar}$ are not arbitrary, but are
given by particular combinations of the
momenta of the $t,\tbar, \ell^+$ and $\ell^-$~\cite{Arens},
%
\bea
\label{eq3}
n_t&=&-\frac{p_t}{m_t}+\frac{m_t}{(p_t\cdot p_{{\ell}^+})}p_{{\ell}^+}\\
\label{eq4}
n_{\tbar}&=&\,\frac{p_{\tbar}}{m_t}-\frac{m_t}{(p_{\tbar}\cdot p_{{\ell}^-})}p_{{\ell}^-}.
\eea
%
Expressions similar to Eq.~(\ref{eq2}) have been derived previously for the
production of short-lived particles in $e^-e^+$ colliders
\cite{kawasaki} and for $t\tbar$ production both in
$e^-e^+$ colliders \cite{Arens} and $pp$ colliders
\cite{ale1,*ale2,*ale3,*ale4}. \par

@@

\subsection{Origin of Triple product terms}

@@

The results shown in Table~\ref{table1}
and Fig.~\ref{fig3} all assume a $pp$ initial state, which is
actually a combination of events coming from $gg$ and $q\qbar$
initial states.
While this combination of initial states is the appropriate scenario
to consider, it
is interesting to consider the relative contributions to the asymmetry
coming from the
$gg$ and $q\qbar$ initial states.  Figure~\ref{fig4} shows
three curves for the ``$\epsilon_1$'' case, one for $gg$-initiated
events, one for $q\qbar$-initiated events, and one for the usual
combination of these events (the``$pp$'' initial state).
Interestingly, we see from Fig.~\ref{fig4} that the asymmetry for this TP
is enhanced for
$gg$-initiated production, while it is reduced and of opposite sign
for the $q\qbar$-initiated events. The asymmetry for the $pp$ case is
evidently
dominated by the $gg$ contribution, but is somewhat smaller
in magnitude due to the
$q\qbar$ contribution.
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\renewcommand{\arraystretch}{1.4}
\begin{table}[H]
\caption{Asymmetries obtained by using $10^5$ simulated events for the
  TPs $\epsilon_1=\TPa, \epsilon_2=\TPb$ and $\epsilon_3=\TPc$ in the
  SM case and the two $\mathrm{CP}$-mixed cases defined by
  $\kp=1,\kpt=\pm 1$.}
\label{table1}
\begin{center}
\begin{tabular}{|C{1cm}|C{1cm}||C{2cm}|C{2cm}||C{2cm}|C{2cm}||C{2cm}|C{2cm}|}
%\begin{tabular}{|c|r||r|c||r|c||r|c|}
\hhline{|========|}
%\hhline{|--------|}
$\kappa_t$&$\tilde{\kappa}_t$~~&$\mathcal{A}(\epsilon_1)$~~&$\mathcal{A}(\epsilon_1)/\sigma_{\mathcal{A}}$& $\mathcal{A}(\epsilon_2)$~~&$\mathcal{A}(\epsilon_2)/\sigma_{\mathcal{A}}$&$\mathcal{A}(\epsilon_3)$~~&$\mathcal{A}(\epsilon_3)/\sigma_{\mathcal{A}}$  \\ 
\hhline{|========|} 
%\hhline{|--------|}
$1$ & $-1$~~~ & $0.0315$~ & $10.0$ & $0.0332$~ & $10.5$~ & $-0.0307$~~~ & $-9.7$~~~\\[0.6mm]
\hline
$1$ & $0$ & $-0.0021$~~~ & $-0.7$~~~ & $0.0009$~ & $0.3$~ & $-0.0011$~~~ & $-0.3$~~~\\[0.6mm]
\hline
$1$ & $1$ & $-0.0379$~~~ & $-12.0$~~~ & $-0.0411$~~~& $-13.0$~~~ & $\,0.0378$~ & $12.0$ \\[0.6mm]
\hhline{|========|}
%\hhline{|--------|}
\end{tabular}
\end{center} 
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% FIGURA 3 %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\vspace*{2.5mm}
\begin{figure}[H]
%\centering
\hspace*{-0.45cm}
\subfloat{\includegraphics[scale=0.45]{ATP1_nuevo.pdf}}
\hspace*{0.002\textwidth}
\subfloat{\includegraphics[scale=0.45]{ATP2_nuevo.pdf}} \\
%\hspace*{0.03\textwidth}
\centering
\subfloat{\includegraphics[scale=0.45]{ATP3_nuevo.pdf}}
\caption{Asymmetries for the TPs
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$ (top-left),
  $\epsilon_2=\epsilon(Q,\tbar,n_t,n_{\tbar})$ (top-right) and
  $\epsilon_3=\epsilon(Q,t,n_t,n_{\tbar})$ (bottom). The points
  represent the values for $\kpt=0,\pm 0.25, \pm 0.5, \pm 0.75,\pm 1$
  and the red solid line is the fitting curve.}
\label{fig3}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%% FIGURA 4  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\vspace*{-4mm}
\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{ATP1juntos_nuevo.pdf}
\caption{Asymmetry for the TP
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$. The dashed line (red)
  corresponds to $gg$-initiated production, the dot-dashed line (grey)
  to $q\qbar$-initiated production and the solid line (blue) to $pp$
  production.}
\label{fig4}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-6mm}

We have also tested various
linear combinations of the TPs $\epsilon_{1,2,3}$ and have found
that the asymmetry is enhanced for the following combination:
\setlength{\abovedisplayskip}{9.5pt}
\setlength{\belowdisplayskip}{9.5pt}
\label{eq24}
\epsilon_4=\epsilon_3-\epsilon_2=\epsilon(Q,t-\tbar,n_t,n_{\tbar}).
Note that in the $Q$ rest frame, $\epsilon_4=Q^0
(\vec{t}-\vec{\tbar}\,)\cdot(\vec{n}_t\times \vec{n}_{\tbar})$ and the
sign of this TP is determined by the quantity
$(\vec{t}-\vec{\tbar}\,)\cdot(\vec{n}_t\times \vec{n}_{\tbar})$.
{\bf (KK: Is the result from the previous sentence used somewhere?)}
The
values obtained for the asymmetry associated with this TP are shown in
Table~\ref{table2}. By comparing the results in Tables \ref{table1}
and \ref{table2}, we see that the capability of this
asymmetry to distinguish between the
two $\mathrm{CP}$-mixed hypotheses is increased by at least $3\sigma$.
\vspace{4mm}
\caption{Asymmetry for the TP $\epsilon_{4}$ for the SM case and the
\label{table2}
$\kappa_t$&$\tilde{\kappa}_t$~~&$\mathcal{A}(\epsilon_4)$~~&$\mathcal{A}(\epsilon_4)/\sigma_{\mathcal{A}}$ \\ 
$1$ & $-1$~~~ & $-0.0371$~~~ & $-12$~~~ \\[0.6mm]
$1$ & $0$ & $0.0004$~ & $0.1$ \\[0.6mm]
$1$ & $1$ & $0.0461$~ & $14\,$ \\[0.6mm]
\par Finally, it is worth noting that the asymmetries
described in this subsection are not useful for
discriminating between the SM hypothesis ($\kp=1,\kpt=0$) and the pure
pseudoscalar hypothesis ($\kp=0,\kpt=1$).  Since
the numerators of the 
asymmetries are all linear in both
$\kp$ and $\kpt$, they expected to vanish in these cases ({\bf ``..they are expected to.."? Maybe it's a typo? we are not sure}). However, we
will show in the next subsection
that there exist angular distributions
derived from the TPs that are actually suitable for distinguishing between these
two hyphothesis.
\subsection{Angular Distributions}
\label{sec3.2}
Given a certain TP, it is possible to define asocciated angular
distributions that are also sensitive to the pseudoscalar coupling

@@

$\epsilon(t+\tbar,\tbar,n_t,n_{\tbar})$ so that in the reference frame

@@

of $\vec{n}_t$ and $\vec{n}_{\tbar}$ respectively and finally

@@

perpendicular to $\vec{\bar{t}}$. If we define this angle within the

@@

determine the sign of the TP. In this sense, the distribution of the
number of the events with respect to the angle

@@

where $N_T$ is the total number of events. Moreover, from a certain TP

@@

different reference frames, but all of these will satisfy
Eq.~(\ref{eq26}). In particular, we will analyze the following angular
distributions based on the TPs defined in Sec.~\ref{sec2}:
\begin{itemize}
\item Related to $\epsilon_1 = \TPa$ we consider
  $d\sigma/d\Delta\phi_1(n_t,n_{\tbar})$ in the rest frame of $t\tbar$
  with $\vec{\tbar}$ in the $z$-axis. The angle

@@

\item Related to $\epsilon_2 = \TPb$ we consider
  $d\sigma/d\Delta\phi_2(n_t,n_{\tbar})$ in the rest frame of $Q$ with
  $\vec{\tbar}$ in the $z$-axis. The angle

@@

\item Related to $\epsilon_3 = \TPc$ we consider
  $d\sigma/d\Delta\phi_3(n_t,n_{\tbar})$ in the rest frame of $Q$ with
  $\vec{t}$ in the $z$-axis. The angle $\Delta\phi_3(n_t,n_{\tbar})$

@@

\end{itemize}

@@

In Fig.~\ref{fig5} we present the normalized distributions obtained
for the first case listed above with $\kp= 1,\kpt=0$
($\mathrm{CP}$-even Higgs boson), $\kp= 1,\kpt=\pm 1$
($\mathrm{CP}$-mixed Higgs boson) and $\kp= 0,\kpt=1$
($\mathrm{CP}$-odd Higgs boson). In Fig.~\ref{fig6} we display in turn

@@

correponding to $\epsilon_3$ are similar to those of $\epsilon_2$ but
with the opposite shift for $\kp=\kpt=1$ and $\kp=-\kpt=1$ and then we
do not include them here.

@@

  $\epsilon(t,\tbar,n_t,n_{\tbar})$ for various values of
  $(\kp,\kpt)$. The error bars correspond to the statistical

@@

\caption{Angular distributions associated with the TP $\TPb$ for
  various values of $(\kp,\kpt)$. The error bars correspond to the
  statistical uncertainties.}

@@

\par As can be seen from Figs.~\ref{fig5} and \ref{fig6}, the peak of
the distributions is shifted according to the values of the couplings
$\kp, \kpt$. This shift appears to be approximately the same but with
opposite sign for $\kp=\kpt=1$ and $\kp=-\kpt=1$, distinguishing then
the sign of the pseudoscalar coupling. The last is the expected
behaviour since the asymmetry is linear in $\kpt$ (see
Sec.~\ref{sec3.1}) and the quantity $N(\epsilon < 0)/N_T$ is related
to the angular distribution according to Eq.~(\ref{eq26}). The
distributions for the SM and the pure pseudoscalar hypotheses are
clearly different. While in the former case the distributions exhibit
a minimum at $\Delta\phi_{1,2}(n_t,n_{\tbar})=0$, in the latter the
similar distributions display a peak, so that the two hypotheses can
be discriminated, in contrast to the asymmetry that vanishes for both
hypotheses. In order to quantify the mentioned shifts, we have fitted

>>>>

At the LHC $t\bar{t}H$ production proceeds via $q\bar{q}$
annihilation and $gg$ fusion processes. The relevant leading-order Feynman
diagrams are displayed in Fig.~\ref{fig1}, where the first two
rows show the $q\bar{q}$ and $gg$ $s$-channel diagrams, and
the last one depicts the $gg$ $t$-channel diagrams.  Three more
$gg$-initiated diagrams are obtained by exchanging the gluon
lines in the third row.
We describe the $tH$ coupling with the effective Lagrangian
%
\beq
\label{eq1}
\mathcal{L}_{t\bar{t}H}=-\frac{m_t}{v}(\kp \tbar t+i\kpt \tbar
\gamma_5 t)H,
\eeq
%
where $v=246~\mathrm{GeV}$ is the SM Higgs vacuum
expectation value, and the coefficients $\kp$ and $\kpt$ parameterize 
the scalar and pseudoscalar interaction, respectively. The
SM case is obtained for $\kp=1$ and $\kpt=0$, while the values $\kp=0$
and $\kpt\neq 0$ parameterize a $\mathrm{CP}$-odd Higgs boson.

Before turning to a discussion of $\mathrm{CP}$-odd observables,
it is useful to consider a few theoretical aspects of
the process $\ppprocess$.
In the following subsections we derive a ``factorized''
expression for the gluon fusion contribution to this process
and then use this expression
to isolate various mathematical quantities
that will be useful as we construct $\mathrm{CP}$-odd observables.\par

%%%%%%%%%%%%%%%%%%%% Figura 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/home/nico/jaxodraw-2.1-0/
\begin{center}
\begin{figure}[H]
\centering
%\hspace*{-0.4cm}
\subfloat{\includegraphics[scale=0.45]{qqs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1a}}
\subfloat{\includegraphics[scale=0.45]{qqs2_II.pdf}}
%\label{fig1b}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1b}}
\subfloat{\includegraphics[scale=0.45]{ggs2_II.pdf}}
%\label{fig1c}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggt1_II.pdf}}
%\label{fig1b}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt2_II.pdf}}
%\label{fig1c}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt3_II.pdf}}
%\label{fig1b}}
\vspace*{0.02\textwidth}
\caption{Tree-level Feynman diagrams contributing to $t\tbar H$ production
  at the LHC. Three more diagrams are obtained by exchanging the gluon
  lines in the $t$-channel diagrams.}
\label{fig1}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Factorized expression for the scattering cross section}
\label{subsec:factorize}

In this subsection we focus on the gluon fusion
{\bf (KK: seems repetitive to say ``gluon fusion'' right away
  again\ldots maybe can change this; maybe OK to leave it?)}
contributions
to $t\tbar H$ production, since these dominate over the
the quark-antiquark annihilation contributions.  Furthermore,
we consider the case in which the top and antitop both decay
semileptonically.  As we shall show below, assuming the
narrow width approximation for the top and antitop quarks, the unpolarized
differential cross section for $gg\to t(\to
b{\ell}^+\nu_{\ell})\,\tbar(\to \bbar {\ell}^- \nubar)\,H$ may be
written in the following ``factorized'' form,\footnote{The reader
  is referred to the discussion following Eq.~(\ref{eq17}) for some qualifying
  remarks regarding the ``factorization'' of this expression.}
%
\beq
\label{eq2}
%dd\sigma(gg\to (bl^+\nu_l)(\bbar l^- \nubar) H)=\sum_{\substack{bl^+\nu_l \\ \tiny{\mathrm{spins}}}}\,\sum_{\substack{\bbar l^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,d\Gamma(t(n_t)\to bl^+\nu_l)\,d\Gamma(\tbar (n_{\tbar})\to \bbar l^-\nubar)
d\sigma =\sum_{\substack{b{\ell}^+\nu_l \\ \tiny{\mathrm{spins}}}}\,
   \sum_{\substack{\bbar {\ell}^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,
   d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,
   d\Gamma(t\to b{\ell}^+\nu_{\ell})\,
   d\Gamma(\tbar \to \bbar {\ell}^-\nubar),
\eeq  
%
where $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$ is the differential
cross section for the production of a top and antitop quark,
with spin vectors $n_t$ and $n_{\bar{t}}$, respectively, along with a Higgs
boson.  Also, $d\Gamma(t\to b{\ell}^+\nu_{\ell})$ and
$d\Gamma(\tbar \to \bbar {\ell}^-\nubar)$ are the
partial differential decay widths for an unpolarized top and
anti-top quark.  The four-vectors
$n_t$ and $n_{\tbar}$ are not arbitrary, but are
given by particular combinations of the
momenta of the $t,\tbar, \ell^+$ and $\ell^-$~\cite{Arens},
%
\bea
\label{eq3}
n_t&=&-\frac{p_t}{m_t}+\frac{m_t}{(p_t\cdot p_{{\ell}^+})}p_{{\ell}^+}\\
\label{eq4}
n_{\tbar}&=&\,\frac{p_{\tbar}}{m_t}-\frac{m_t}{(p_{\tbar}\cdot p_{{\ell}^-})}p_{{\ell}^-}.
\eea
%
Expressions similar to Eq.~(\ref{eq2}) have been derived previously for the
production of short-lived particles in $e^-e^+$ colliders
\cite{kawasaki} and for $t\tbar$ production both in
$e^-e^+$ colliders \cite{Arens} and $pp$ colliders
\cite{ale1,*ale2,*ale3,*ale4}. \par

@@

\subsection{Origin of triple product terms}
\label{subsec:origin}

@@

The results shown in Table~\ref{table1}
and Fig.~\ref{fig3} all assume a $pp$ initial state, which is
actually a combination of events coming from $gg$ and $q\qbar$
initial states.
While this combination of initial states is the appropriate scenario
to consider, it
is interesting to consider the relative contributions to the asymmetry
coming from the
$gg$ and $q\qbar$ initial states.  Figure~\ref{fig4} shows
three curves for the ``$\epsilon_1$'' case, one for $gg$-initiated
events, one for $q\qbar$-initiated events, and one for the usual
combination of these events (the ``$pp$'' initial state).
Interestingly, we see from Fig.~\ref{fig4} that the asymmetry for this TP
is enhanced for
$gg$-initiated production, while it is reduced and of opposite sign
for the $q\qbar$-initiated events. The asymmetry for the $pp$ case is
evidently
dominated by the $gg$ contribution, but is somewhat smaller
in magnitude due to the
$q\qbar$ contribution.
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\renewcommand{\arraystretch}{1.4}
\begin{table}[H]
  \caption{Asymmetries for three different scenarios,
    obtained by using $10^5$ simulated events for the
    TPs $\epsilon_1=\TPa, \epsilon_2=\TPb$ and $\epsilon_3=\TPc$.
    The three scenarios
  correspond to the SM ($\kp=1$ and $\kpt=\pm 0$) and
  two $\mathrm{CP}$-mixed cases (defined by
  $\kp=1$ and $\kpt=\pm 1$).}
\label{table1}
\begin{center}
\begin{tabular}{|C{1cm}|C{1cm}||C{2cm}|C{2cm}||C{2cm}|C{2cm}||C{2cm}|C{2cm}|}
%\begin{tabular}{|c|r||r|c||r|c||r|c|}
\hhline{|========|}
%\hhline{|--------|}
$\kappa_t$&$\tilde{\kappa}_t$~~&$\mathcal{A}(\epsilon_1)$~~&$\mathcal{A}(\epsilon_1)/\sigma_{\mathcal{A}}$& $\mathcal{A}(\epsilon_2)$~~&$\mathcal{A}(\epsilon_2)/\sigma_{\mathcal{A}}$&$\mathcal{A}(\epsilon_3)$~~&$\mathcal{A}(\epsilon_3)/\sigma_{\mathcal{A}}$  \\ 
\hhline{|========|} 
%\hhline{|--------|}
$1$ & $-1$~~~ & $0.0315$~ & $10.0$ & $0.0332$~ & $10.5$~ & $-0.0307$~~~ & $-9.7$~~~\\[0.6mm]
\hline
$1$ & $0$ & $-0.0021$~~~ & $-0.7$~~~ & $0.0009$~ & $0.3$~ & $-0.0011$~~~ & $-0.3$~~~\\[0.6mm]
\hline
$1$ & $1$ & $-0.0379$~~~ & $-12.0$~~~ & $-0.0411$~~~& $-13.0$~~~ & $\,0.0378$~ & $12.0$ \\[0.6mm]
\hhline{|========|}
%\hhline{|--------|}
\end{tabular}
\end{center} 
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% FIGURA 3 %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\vspace*{2.5mm}
\begin{figure}[H]
%\centering
\hspace*{-0.45cm}
\subfloat{\includegraphics[scale=0.45]{ATP1_nuevo.pdf}}
\hspace*{0.002\textwidth}
\subfloat{\includegraphics[scale=0.45]{ATP2_nuevo.pdf}} \\
%\hspace*{0.03\textwidth}
\centering
\subfloat{\includegraphics[scale=0.45]{ATP3_nuevo.pdf}}
\caption{Asymmetries for the TPs
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$ (top-left),
  $\epsilon_2=\epsilon(Q,\tbar,n_t,n_{\tbar})$ (top-right) and
  $\epsilon_3=\epsilon(Q,t,n_t,n_{\tbar})$ (bottom). The points
  represent the values for $\kpt=0,\pm 0.25, \pm 0.5, \pm 0.75,\pm 1$
  and the red solid line is the fitting curve.}
\label{fig3}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%% FIGURA 4  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\vspace*{-4mm}
\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{ATP1juntos_nuevo.pdf}
\caption{Asymmetry for the TP
  $\epsilon_1=\epsilon(t,\tbar,n_t,n_{\tbar})$. The dashed line (red)
  corresponds to $gg$-initiated production, the dot-dashed line (grey)
  to $q\qbar$-initiated production and the solid line (blue) to $pp$
  production.}
\label{fig4}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-6mm}

We have also tested various
linear combinations of the TPs $\epsilon_{1,2,3}$ and have found
that the asymmetry is enhanced for the following combination:
%
\beq
\setlength{\abovedisplayskip}{9.5pt}
\setlength{\belowdisplayskip}{9.5pt}
\label{eq24}
\epsilon_4=\epsilon_3-\epsilon_2=\epsilon(Q,t-\tbar,n_t,n_{\tbar}).
\eeq
%
Note that in the $Q$ rest frame, $\epsilon_4=Q^0
(\vec{t}-\vec{\tbar}\,)\cdot(\vec{n}_t\times \vec{n}_{\tbar})$ and the
sign of this TP is determined by the quantity
$(\vec{t}-\vec{\tbar}\,)\cdot(\vec{n}_t\times \vec{n}_{\tbar})$.
{\bf (KK: try to connect previous sentences to something later on\ldots.)}
The
values obtained for the asymmetry associated with this TP are shown in
Table~\ref{table2}. By comparing the results in Tables \ref{table1}
and \ref{table2}, we see that the capability of this
asymmetry to distinguish between the
two $\mathrm{CP}$-mixed hypotheses is increased by at least $3\sigma$.
\vspace{4mm}
\begin{table}[H]
\caption{Asymmetry for the TP $\epsilon_{4}$ for the SM case and the
  two $\mathrm{CP}$-mixed cases defined by $\kp=1,\kpt=\pm 1$. The
  values are obtained by using $10^5$ simulated events.}
\label{table2}
\begin{center}
\begin{tabular}{|C{1cm}|C{1cm}||C{2cm}|C{2cm}|}
%\begin{tabular}{|c|r||r|c||r|c||r|c|}
\hhline{|====|}
%\hhline{|--------|}
$\kappa_t$&$\tilde{\kappa}_t$~~&$\mathcal{A}(\epsilon_4)$~~&$\mathcal{A}(\epsilon_4)/\sigma_{\mathcal{A}}$ \\ 
\hhline{|====|} 
%\hhline{|--------|}
$1$ & $-1$~~~ & $-0.0371$~~~ & $-12$~~~ \\[0.6mm]
\hline
$1$ & $0$ & $0.0004$~ & $0.1$ \\[0.6mm]
\hline
$1$ & $1$ & $0.0461$~ & $14\,$ \\[0.6mm]
\hhline{|====|}
%\hhline{|--------|}
\end{tabular}
\end{center} 
\end{table}
\par Finally, it is worth noting that the asymmetries
described in this subsection are not useful for
discriminating between the SM hypothesis ($\kp=1,\kpt=0$) and the pure
pseudoscalar hypothesis ($\kp=0,\kpt=1$).  Since
the numerators of the 
asymmetries are all linear in both
$\kp$ and $\kpt$, they are expected to vanish in these cases. However, we
will show in the next subsection
that there exist angular distributions
derived from the TPs that are actually suitable for distinguishing between these
two hyphotheses.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Angular Distributions}
\label{sec3.2}

Given a certain TP, it is possible to define associated angular
distributions that are sensitive to the pseudoscalar coupling

@@

$\epsilon(t+\tbar,\tbar,n_t,n_{\tbar})$, so that in the reference frame

@@

of $\vec{n}_t$ and $\vec{n}_{\tbar}$, respectively, and 

@@

perpendicular to $\vec{\bar{t}}$. If we define the angle
$\Delta\phi(n_t,n_{\tbar})$  to be within the

@@

determine the sign of the TP. Thus, the distribution of the
number of events with respect to the angle

@@

where $N_T$ is the total number of events.
{\bf (KK: Have we checked numerically that the above
  expression does indeed hold numerically for the various
  distributions?  It's an exact relation, correct?)}
Moreover, for a certain TP

@@

different reference frames, although all of these will satisfy
Eq.~(\ref{eq26}) (note that $\mathcal{A}(\epsilon)$
is Lorentz invariant). Recalling the various TPs considered
in Sec.~\ref{sec2}, we examine the following angular distributions.
\begin{enumerate}
\item {\boldmath $\epsilon_1 = \TPa$.}  To probe $\epsilon_1$, we construct
  the distribution
  $d\sigma/d\Delta\phi_1(n_t,n_{\tbar})$ in the rest frame of $t\tbar$,
  taking $\vec{\tbar}$ to define the $z$-axis. The angle

@@

\item {\boldmath $\epsilon_2 = \TPb$.}  In this case, we
  define the distribution
  $d\sigma/d\Delta\phi_2(n_t,n_{\tbar})$ in the rest frame of $Q$, taking
  $\vec{\tbar}$ to define the $z$-axis. The angle

@@

\item {\boldmath $\epsilon_3 = \TPc$.}  The distribution
  $d\sigma/d\Delta\phi_3(n_t,n_{\tbar})$ is also defined
  in the rest frame of $Q$, but this time taking
  $\vec{t}$ to be along the $z$-axis. The angle $\Delta\phi_3(n_t,n_{\tbar})$

@@

\end{enumerate}

@@

Figure~\ref{fig5} shows the normalized distributions obtained
for the first case listed above.
Four scenarios are considered, corresponding to the SM
($\kp= 1$ and $\kpt=0$), two cases in which the Higgs boson
has mixed $\mathrm{CP}$ couplings ($\kp= 1$ and $\kpt=\pm 1$)
and a case in which the Higgs boson is purely 
$\mathrm{CP}$-odd ($\kp= 0,\kpt=1$). Figure~\ref{fig6} shows

@@

corresponding to $\epsilon_3$ are similar to those of $\epsilon_2$,
except that the ``shifts'' are in the opposite directions for
the two mixed-$\mathrm{CP}$ cases.  Given the similarities of the
plots we do not include them here.  {\bf (KK: did I reword this correctly?)}

@@

  $\epsilon_1 = \epsilon(t,\tbar,n_t,n_{\tbar})$ for various values of
  $\kp$ and $\kpt$. The error bars correspond to the statistical

@@

\caption{Angular distributions associated with the TP $\epsilon_2 = \TPb$ for
  various values of $\kp$ and $\kpt$. The error bars indicate the statistical

@@

\par As can be seen from Figs.~\ref{fig5} and \ref{fig6}, the
peaks of the 
distributions are shifted to the left or the right of the origin
in the mixed-$\mathrm{CP}$
cases ($\kp =1$ and $\kpt=\pm 1$).
The magnitude of the shift appears to be approximately the
same in both cases, but with
opposite sign for $\kp=\kpt=1$ and $\kp=-\kpt=1$, thus allowing
one to distinguish
the sign of the pseudoscalar coupling. The observed dependence
on the sign of $\kpt$ in these cases
is consistent with the fact that the numerator of $\mathcal{A}(\epsilon)$
is linear in $\kpt$ [see Eq.~(\ref{eq26})]
and that the quantity $N(\epsilon < 0)/N_T$ is related
to the angular distribution according to Eq.~(\ref{eq26}). The angular
distributions for the SM case ($\kp =1$ and $\kpt=0$) and
the pure pseudoscalar case ($\kp =0$ and $\kpt= 1$) are
visibly different from each other and from the mixed-$\mathrm{CP}$
scenarios. Comparing the SM and purely pseudoscalar cases,
we note that while the angular distributions for the former case
